### 2019-03-11 paper.tex
Added online phase/ only thing missing is the improvemetn with different beaver triples.

% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.20 (10-Mar-2018)

% !TeX spellcheck = en-US
% !TeX encoding = utf8
% !TeX program = pdflatex
% !BIB program = bibtex
% -*- coding:utf-8 mod:LaTeX -*-

% "a4paper" enables:
%  - easy print out on DIN A4 paper size
%
% One can configure a4 vs. letter in the LaTeX installation. So it is configuration dependend, what the paper size will be.
% This option  present, because the current word template offered by Springer is DIN A4.
% We accept that DIN A4 cause WTFs at persons not used to A4 in USA.

% "runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
% This is good for other readers to enable proper archiving among other papers and pointing to
% content. Even if the title page states the title, when printed and stored in a folder, when
% blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
% it is good to have title printed on the pages, too.
%
% It is enabled by default as the springer template as of 2018/03/10 uses this as default

% German documents: pass ngerman as class option
% \documentclass[ngerman,runningheads,a4paper]{llncs}[2018/03/10]
% English documents: pass english as class option
\documentclass[english,runningheads,a4paper]{llncs}[2018/03/10]

%% If you need packages for other papers,
%% START COPYING HERE

% Set English as language and allow to write hyphenated"=words
%
% In case you write German, switch the parameters, so that the command becomes
%\usepackage[english,main=ngerman]{babel}
%
% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
% english has to go last to set it as default language
\usepackage[ngerman,main=english]{babel}
%
% Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}
%
% Fix by https://tex.stackexchange.com/a/441701/9075
\usepackage{regexpatch}
\makeatletter
\edef\switcht@albion{%
  \relax\unexpanded\expandafter{\switcht@albion}%
}
\xpatchcmd*{\switcht@albion}{ \def}{\def}{}{}
\xpatchcmd{\switcht@albion}{\relax}{}{}{}
\edef\switcht@deutsch{%
  \relax\unexpanded\expandafter{\switcht@deutsch}%
}
\xpatchcmd*{\switcht@deutsch}{ \def}{\def}{}{}
\xpatchcmd{\switcht@deutsch}{\relax}{}{}{}
\edef\switcht@francais{%
  \relax\unexpanded\expandafter{\switcht@francais}%
}
\xpatchcmd*{\switcht@francais}{ \def}{\def}{}{}
\xpatchcmd{\switcht@francais}{\relax}{}{}{}
\makeatother

\usepackage{url}
\usepackage{ifluatex}
\ifluatex
  \usepackage{fontspec}
  \usepackage[english]{selnolig}
\fi

\iftrue % use default-font
  \ifluatex
    % use the better (sharper, ...) Latin Modern variant of Computer Modern
    \setmainfont{Latin Modern Roman}
    \setsansfont{Latin Modern Sans}
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % better font, similar to the default springer font
    % cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
    \usepackage[%
      rm={oldstyle=false,proportional=true},%
      sf={oldstyle=false,proportional=true},%
      tt={oldstyle=false,proportional=true,variable=false},%
      qt=false%
    ]{cfr-lm}
  \fi
\else
  % In case more space is needed, it is accepted to use Times New Roman
  \ifluatex
    \setmainfont{TeX Gyre Termes}
    \setsansfont[Scale=.9]{TeX Gyre Heros}
    % newtxtt looks good with times, but no equivalent for lualatex found,
    % therefore tried to replace with inconsolata.
    % However, inconsolata does not look good in the context of LNCS ...
    %\setmonofont[StylisticSet={1,3},Scale=.9]{inconsolata}
    % ... thus, we use the good old Latin Modern Mono font for source code.
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % overwrite cmodern with the Times variant
    \usepackage{newtxtext}
    \usepackage{newtxmath}
    \usepackage[zerostyle=b,scaled=.9]{newtxtt}
  \fi
\fi

\ifluatex
\else
  % fontenc and inputenc are not required when using lualatex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc} %support umlauts in the input
\fi

\usepackage{graphicx}

% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
\usepackage{upquote}

% Nicer tables (\toprule, \midrule, \bottomrule - see example)
\usepackage{booktabs}

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

% For easy quotations: \enquote{text}
% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
\usepackage{csquotes}

% For even easier quotations: \qq{text}
\usepackage{textcmds}

%enable margin kerning
\RequirePackage[%
  babel,%
  final,%
  expansion=alltext,%
  protrusion=alltext-nott]{microtype}%
% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
\DisableLigatures{encoding = T1, family = tt* }

%tweak \url{...}
\usepackage{url}
%\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
%DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
% Slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

% Required for package pdfcomment later
\usepackage{xcolor}

% For listings
\usepackage{listings}
\lstset{%
  basicstyle=\ttfamily,%
  columns=fixed,%
  basewidth=.5em,%
  xleftmargin=0.5cm,%
  captionpos=b}%
\renewcommand{\lstlistingname}{List.}
% Fix counter as described at https://tex.stackexchange.com/a/28334/9075
\usepackage{chngcntr}
\AtBeginDocument{\counterwithout{lstlisting}{section}}

% Enable nice comments
\usepackage{pdfcomment}
%
\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
%
% Compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}
% Compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

% Bibliopgraphy enhancements
%  - enable \cite[prenote][]{ref}
%  - enable \cite{ref1,ref2}
% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")

% Doc: http://texdoc.net/natbib
\usepackage[%
  square,        % for square brackets
  comma,         % use commas as separators
  numbers,       % for numerical citations;
%  sort,          % orders multiple citations into the sequence in which they appear in the list of references;
  sort&compress, % as sort but in addition multiple numerical citations
                 % are compressed if possible (as 3-6, 15);
]{natbib}
% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
\renewcommand{\bibnumfmt}[1]{#1.}

\ifluatex
  % does not work when using luatex
  % see: https://tex.stackexchange.com/q/419288/9075
\else
  % Prepare more space-saving rendering of the bibliography
  % Source: https://tex.stackexchange.com/a/280936/9075
  \SetExpansion
  [ context = sloppy,
    stretch = 30,
    shrink = 60,
    step = 5 ]
  { encoding = {OT1,T1,TS1} }
  { }
\fi

% Put footnotes below floats
% Source: https://tex.stackexchange.com/a/32993/9075
\usepackage{stfloats}
\fnbelowfloat

% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
\usepackage{hyperref}
% Enable hyperref without colors and without bookmarks
\hypersetup{hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true}
%
% Enable correct jumping to figures when referencing
\usepackage[all]{hypcap}

\usepackage[group-four-digits,per-mode=fraction]{siunitx}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\usepackage{iflang}
\IfLanguageName{ngerman}{
  \crefname{table}{Tab.}{Tab.}
  \Crefname{table}{Tabelle}{Tabellen}
  \crefname{figure}{\figurename}{\figurename}
  \Crefname{figure}{Abbildungen}{Abbildungen}
  \crefname{equation}{Gleichung}{Gleichungen}
  \Crefname{equation}{Gleichung}{Gleichungen}
  \crefname{listing}{\lstlistingname}{\lstlistingname}
  \Crefname{listing}{Listing}{Listings}
  \crefname{section}{Abschnitt}{Abschnitte}
  \Crefname{section}{Abschnitt}{Abschnitte}
  \crefname{paragraph}{Abschnitt}{Abschnitte}
  \Crefname{paragraph}{Abschnitt}{Abschnitte}
  \crefname{subparagraph}{Abschnitt}{Abschnitte}
  \Crefname{subparagraph}{Abschnitt}{Abschnitte}
}{
  \crefname{section}{Sect.}{Sect.}
  \Crefname{section}{Section}{Sections}
  \crefname{listing}{\lstlistingname}{\lstlistingname}
  \Crefname{listing}{Listing}{Listings}
}

%Intermediate solution for hyperlinked refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
  <-6>    MnSymbolC5
  <6-7>   MnSymbolC6
  <7-8>   MnSymbolC7
  <8-9>   MnSymbolC8
  <9-10>  MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

\ifluatex
\else
  % Enable copy and paste - also of numbers
  % This has to be done instead of \usepackage{cmap}, because it does not work together with cfr-lm.
  % See: https://tex.stackexchange.com/a/430599/9075
  \input glyphtounicode
  \pdfgentounicode=1
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% END COPYING HERE

% Add copyright
% Do that for the final version or if you send it to colleagues
\iffalse
  %state: intended|submitted|llncs
  %you can add "crop" if the paper should be cropped to the format Springer is publishing
  \usepackage[intended]{llncsconf}

  \conference{name of the conference}

  %in case of "llncs" (final version!)
  %example: llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
  \llncs{book editors and title}{0042} %% 0042 is the start page
\fi

% For demonstration purposes only
\usepackage[math]{blindtext}
\usepackage{mwe}

\begin{document}

\title{Secure Multi-Party Computation with SCALE-MAMBA}
%If Title is too long, use \titlerunning
%\titlerunning{Short Title}

%Single insitute
\author{Jonas Rülfing}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
\institute{Seminar: Privacy and Big Data - RWTH Aachen}

%% Multiple insitutes - ALTERNATIVE to the above
% \author{%
%     Firstname Lastname\inst{1} \and
%     Firstname Lastname\inst{2}
% }
%
%If there are too many authors, use \authorrunning
%  \authorrunning{First Author et al.}
%
%  \institute{
%      Insitute 1\\
%      \email{...}\and
%      Insitute 2\\
%      \email{...}
%}

\maketitle

\begin{abstract}
  This thesis takes a look at SMPC especally using SCALE-MAMBA. For that it introduces first what SMPC is,  how SPDZ the precessor of SCALE-MAMBA works and what are the differences. It discusses the functionalities of SCALE-MAMBA, especially the split into online and offline phase, how the beaver tripels are used for sacrificing and in which complexity classes the calculations are done. Finally it compares scale mamba shortly to other competitors in the market and gives an outlook what next technologies are coming right now and are coming in the near future in SMPC.
\end{abstract}

\section{Motivation and Introduction}\label{sec:intromoti}

In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, this might be because there are different parties involved, but also because the amount of data is getting so big that it becomes impossible to have a single entity that processes all the data at once. Furthermore, in a lot of these cases the parties participating in these computations don't want to reveal their own information, because its sensitive or because they don't want to share information with a lot of business value.\\
A basic example is an anonymized voting. In most elections there is a third party that has to aggregate and count the votings made by the votings. For instance in germanys elections this is done by specific people counting a small group of votes and then forwarding the information. The other problem is that it has to be guaranteed that a vote is valid, meaning that the voter has the right to vote and is only voting one time. So if there would be a malicious person in the system, it would be easy for that person to figure out what vote a specific person made.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and shoudn't be publiced under n circumstances. But at the same time a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making diagnosis for new patients it is helpful to compare the symptoms to the diagnosis of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is how to evaluate a common function in which multiple parties have different inputs to this function, without revealing the different inputs of the parties. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not perfmorming on the desired level. So in the last years the research shifted more in the area on how to make the existant solutions performant enough while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the KU Leuven, which is one of the frontrunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short history of SMPC and will try to explain the basic technologies of SMPC to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparision to other SMPC systems which use different approaches, especially considering the usability in different use cases. Ultimately, it will try to give an outlook what might be coming in the next years.\\

\section{Cryptography Basics}\label{sec:basics}

Before diving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader. These basics are not solely used by SMPC, but nevertheless, play a huge role in a lot of protocols.

\subsection{Message Authentification Codes}
Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message. Basically, a MAC is a tag that is generated for a message and that, if it is correct, verifies the identity of the sender and that the content of the message was not changed.\\
For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
\begin{itemize}
\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) or reject(\(0\))\\
\end{itemize}
The correctness can be easily seen: \(\forall K: \text{VER}(K,m,\text{MAC}(K,m))=1\)\\

In SCALE-MAMBA really simple MACs are used. We assume that we have a MAC key \( \alpha \). Then the MAC algorithm used is following:\\
$$\text{MAC}(\alpha,m)\rightarrow \alpha * m$$ 
$$\text{VER}(\alpha,m,t)\rightarrow (1 \Leftrightarrow t = \alpha*m) $$

These algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is possible to get the original message with just having the key and the tag. This is a property that is used for an advantage later on in the protocol.\\

\subsection{Homomorphic Encryption}
Encryption as a basic concept is simple. Similiar to MACs, a tupel (E,D) of algorithms is needed.\\
\begin{itemize}
\item The enncryption algorithm \(\text{E}(K,m)\rightarrow c\) takes a key \(K\) and a message \(m\) and outputs a cyphertext \(t\) that is the encryption of the message \(m\)\\
\item The decryption algorithm \(\text{D}(K,c)\rightarrow m\) takes a key \(K\) and a cyphertext \(m\) and outputs the decrypted message \(m\), assuming c was not modified\\
\end{itemize}
Alternatively, in a public key setting, the key for encryption is public and there is a different key for decryption that is privat.\\
The problem is, that modifying the cypertext in any way makes the decryption useless. In some cases it is desirable do do computations on cypertexts, because then it is possible to make calculations without revealing the unencrypted data. That is why Homomorphic Encryption was developed. Let's assume that we have a ring \((R,+,*)\) as our message space. Fully Homomorphic Encryption (FHE) is preserving the ring structure. Therefore, it is possible to do additions and multiplications of cypertexts and when decrypting the result it is the same value as when doing the same additions and multiplications on the real values.\\
Althought FHE has amazing properties for a lot of usecases, in praxis it is way too slow. Because of all the overhead generated by the encryption scheme to guarantee the preservation of the ring structure, the current algorithms are not feasible for most applications.

\subsection{Somewhat Homomorphic Encryption}
That is where Somewhat Homomorphic Encryption (SHE) becomes relevant. SHE only supports one operation, and in most cases that is addition. This takes away a lot of possibilities but, therefore, the performance gain is immense. So immense, that SHE is used a lot in practice, for example in SCALE-MAMBA.\\
The SHE scheme used in SCALE-MAMBA, is based on the so called Brakerski-Gentry-Vaikuntanathan cryptosystem(BGV).\\

\subsection{Adversary types}
Cryptography is the science of protecting information or communication from an adversary. This adversary is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest. In the are of SMPC we we normally divide between two types of adversaries. In SMPC we expect that everyone but ourselves can be controlled by an adversary.  That means the adversary can control between one and all of the other parties with which we are communicating. The difference to the normal cryptographic adversary model is, that the adversary is one of the parties we are communicating with but not a third party that tries to intercept communication.
\subsubsection{Honest but curious}
This type of adversary tries to gain as much information as possible. But this adversary is obeying to the rules of our protocol, that means it is impossible to know that this party is controlled by a third entity. This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.

\subsection{Malicious adversary}
While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol. This type of adverary can send different inputs to manipulate the results he receives. Furthermore, he wants to interrupt the whole protocol. That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications

It is important to decide between these two types of adversary, because it shows the different securities SMPC has to guarantee. It has to guarantee that no information is unwillingly shared with an illegal party and at the same time verify that the final result has a correct value. Most SMPC protocols try to guarantee safety versus both type of adveraries, bvut there are also approaches that only use passive security, i.e. only protecting versus non malicious adversaries.

\section{SMPC Basics and History}\label{sec:smpc}
\subsection{Beginnings}
As we saw in the introduction, there are a lot current and much more upcoming usecases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret. The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure noone is cheating) and Yaos millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth). \newline
These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
 Now we will take a short look on the two most famous of these protocols, which both were the basis for a whole family of algorithms established in the years to follow until today.

\subsection{Yaos Garbled Circuit}
There are typically two groups of SMPC protocols, in which most of the protocols can be categorized. One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}. The basic idea is, that, in a two party setting, one party, Alice, garbles the boolean circuit that has to be evaluated and sends it to the second party, Bob. Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an ecnrypted way. At the end they exchange the results and can compute the result this way.

\subsection{Shamirs Secret Sharing Scheme}
This paper will not focus more on the family of garbled circuit protocols, but rather on the second family which are based on a so called Linear Secret Sharing Scheme(LSSS). 
These type of protocols were introduced by Adi Shamir in a similiar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties. This is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial. Computations are executed on these shares, and after the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets. This is done by using Lagrange interpolation.\\
 There is a huge family of protocols build on this approach, and it is still evolving every day. This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the SPDZ like protocols.\\
In SMPC, normally all operations are executed in a finite field \(F_p = \{ 0,...,p-1\}\) with a prime p. That means all operations use modulo p at the end on the result to keep the result in the field. Therefore, for now on \(F_p\) will be used and when using an exact application, then p has to be specified depending on the usecase.\\

\subsection{Terminology}
Over the years different terminologies came up in the are of SMPC. Following will be a short introduction into some of them, so their definitions are clear.

\subsubsection{Oblivious Transfer}

\subsubsection{Access structures}
There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved. it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and for runtime complexity. SMPC should run as fast as possible, while still providing all the neccesary security guarantees.

To describe which party has which rights, and how many parties can be corrupted so the security can still be guaranteed, access structures were introduced.
An access structure describes how many parties are needed to form a so called qualified set. A qualified set is a group of parties which can pool their information to extract secret information..  An access structure is called \(Q_l \) if at least l+1 parties are needed to  form a qualified set. 
An (n,t)-threshold scheme with t<n/l means that at least t+1 parties are needed to form a qualified set. Here, n is the number of total parties, and t is number of possible adversaries while not revealing any secrets.
A full threshold scheme means that all parties are needed to form a qualified set. This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret. Of course this has a huge impact on overhead and performance, but in some cases it can give needed security guarantees.
An access structure also doesn't have to be symmetric. In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.

TODO: SOUNDNESS SLACK
TODO: LATTICE BASED CRYPTOGRAPHY

Scale-Mamba is an implementation of SPDZ among other protocols. SPDZ is right now the most efficient protocol in use.

Most LSSS-based MPC protocols split the computation into two parts: an of-

ine phase, in which parties interact using expensive" public-key cryptography
to lay the groundwork for an online phase in which only cheap" informationtheoretic
primitives are required. The online phase is where the actual circuit
evaluation takes place. For the access structures considered in this work, namely
Q2 structures, the offline phase is almost as fast as the online phase. Thus the
goal here is to minimize the cost of communication in both phases.

\section{SCALE-MAMBA}\label{sec:scalemamba}

\subsubsection{Overview}

In the last years, the research in SMPC is constantly ongoing, and there are new approaches or improvements to old approaches appearing regulary. But most of these publications only focus on a small area of the whole picture. SCALE-MAMBA\cite{ScaleMambaDocu} is a project that tries to combine the state-of-the-art research into one system. That means, that it is offering kind of an end-to-end system for SMPC.
It is a research system that is continously in development, therefore, there is not a single final version that can be examined. It offers as well a wide range of protocols for different usecases, i.e. it is possible to use the system in different confidurations. That also means for this thesis, that there is not a single process that can be examined. Therefore, this thesis tries to give more an overview how the process works in general, which are the main, especially which are newest and best, parts and where are the problems and bottlenecks.\\
SCALE-MAMBA consists of two parts. The one part is called Secure Computation Algorithms from LEuven (SCALE) and is the SMPC part. It contains all the computational logic and algorithms for multi-party computations. The second part is called Multiparty AlgorithMs Basic Argot (MAMBA) and is a compiler for the SMPC programs.\\

In general, the SMPC protocol is split in two parts, first an offline phase and secondly and online phase. In the online phase only the final evaluation of the function is executed. In the offline phase, prior to the online phase, preparations are done by all the parties to prepare for the evaluation. This includes already calculating values for possible multiplications and MACs, so that the online phase can be executed faster. This phase is not completly offline, because the parties still exchange some information which are later needed for the evaluation. But this phasae is not that time critically, that means it can be already executed a lot earlier than the online phase, so that, when the function evaluation is really needed, the result can be calculated in a rather fast time compared to a system that still has to set up everything. In SCALE-MAMBA, compared to its predecessors, the offline and online phase are completly integrated. Therefore, it is not possible anymore to execute them separately like in SPDZ for example. But it is a much closer approach to a real world application.\\

While SCALE-MAMBA also offers different access structures, the most commonly used security configuration is active security with abort in a full threshold scheme, that means that every except one party could be corrupted, and still no information is leaked, because the evaluation is aborted.

\subsubsection{MAMBA}
For this paper the more interesting part is the algorithmic part. But still, the importance of the compiler shouldn't be undervalued. Because while the algorthmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to really use these technologies in a system where he can rely on the security gurantees given by the algorithms. The programs can be written for MAMBA in a syntax similiar to python's. In these programs, the user can specify which security settings he wishes to use. He can specify, how he wants his SMPC to exactly be executed, and which security parameters he wants to use. MAMBA is the compiler which then translates the instruction to bytecode and executes them. MAMBA is the part that gives the system the possibility to be properly used and tested. It makes it easy, to change the parameters, to be able to execute comparable tests, and to have a abstraction level for the user.\\

\subsubsection{Offline phase}

In the offline or pre-processing phase, the most important part is the generation of the beaver triples. These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
Beaver triples are a tuple of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F} )\), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).

The second thing that is done in the offline phase is the generation of a global MAC key. This MAC key is later used, to authentificate the different shares at the end of the computation. This key is also distributedly generated, i.e. only if all parties pool together they can recover its value.\\

\subsubsection{Zero Proofs of Knowledge}
In the many iterations of the SPDZ like protocols, there were different ways to guarantee that the parties are not cheating. An important technology for that are the so called Zero Proofs of Konwledge Protocol(ZPoKP).

In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly, or atleast that they dont divert more than a specified error \(e\) the expected result.

ZPoKP are a technique that was introduced  in 1985 \cite{Goldwasser:1985:KCI:22145.22178}. The idea is that one party can prove to another party that it knows a specific value \(x\), without revealing the value of \(x\) to the other person, nor revealing any other knowledge about the value.

Assumed Alice(in this scenario called the prover) wants to prove to Bob (the verifier) that she knows \( x\) in a Field \( \mathcal{F} \), such that \(f(x)=y\). It is as well assumed, that \( f \) is homomorphic with respect to the field operations in  \( \mathcal{F} \).
\begin{enumerate}
\item Alice chooses a random \( s \in \mathcal{F} \) , calculates \( a = f(s)\) and sends \( a\) to Bob
\item Bob samples a random  \( e \) out of  \( \mathcal{F} \) and sends it to Alice
\item Alice calculates \( z=s+e*x\) and sends it to Bob
\item Bob checks if \( f(z)=a+e*y\)
\end{enumerate}

\textbf{Correctness:}
Bob wants to prove that \( z\) provided by Alice contains the proper \( x\), so that \( f(x)=y\). 
Because f is homomorphic, obviously \( f(z)=f(s)+f(e*x)=f(s)+e*f(x)=a+e*y\) with \(f(s)=a\) and \(f(x)=y\).
So Bob has to calculate \( a+e*y and f(z)\) with the provided \( z\) by Alice. If they are equal, he knows that Alice knows the \( x\) so that \( f(x)=y\).
He can do that, because he got provided the \( a\) and \( z\) by Alice, he knew the \( y\) before the protocol and he chose the \( e\) himself.\\

\textbf{Confidentiality:}
We also have to make sure, that Alice did not share any information regarding her x.
In the firt step she chose a random element s and calulated a=f(s). This information is independent on her secret x, so she didn;t share any information.
When calculating z=s+e*x, she also didnt share any information regarding x, as long it is guaranteed that Bob didnt have any knowledge about the s.
So under the assumption that Alice chooses a new random s each time the protcol is executed, she doesn't share any information regarding her secret x.\\

It is not really important for this thesis, which ZPoKP is exactly used, but it should be said that the protocol used in SCALE-MAMBA is based on Schnorr's protocol.\\

\subsubsection{First improvement with MASCOT}
In one of there earlier works, BDOZ, pairwise MACs were used to authentificate the secret sharing between the parties. It used a pairwise multiplication protocol with linear homomorpic encryption. To guarantee the data integritym pairwise zero-knowledge proofs were used, in total O(n2) of them. 
In SPDZ instead of HE, SHE based on the BGV ecnryption scheme was used, reducing the amount of ZK proofs per triple by a factor of n and as well the overhead per triple, because they used a single ZKPoK for proving thousand of triples.
Because the offline phase was the weak point of the existing SPDZ implementations, in regards to runtime while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2. MASCOT used Oblivious Transfer instead of SHE to significantly reduce the amount of communication and computation. We will not go into detail into this protocol, but rather look on its successor.

\subsubsection{Overdrive}
Shortly afterwards in 2017 a new protocol for the offline phase called Overdrive was introduced \cite{cryptoeprint:2017:1230}, which once again got better results with SHE instead of OT, improving the run time at least by a factor of 6. Basically, it is a really similiar protocol to MASCOT, but just uses SHE in the places where MASCOT used OT. So explaining Overdrive already should give a good impression how MASCOT works. \\
 Overdrive includes two parts, a protocol called Low Gear protocol for small numbers, and a proctocol called High Gear for larger numbers. In version 1.2 of Scale-mamba only the HighGear protocol was used, even for smaller numbers.\\

The biggest difference between the ZKPoK usage in Overdrive and in SPDZ is, that the parties not anymore prove multiple smaller statements between each other with ZKPoK, but instead make a joint statement and prove thi one toegether. Because ZKPoK are really expensive, this makes a big difference in the overall performance.\\
This way, Overdrive did not improve the amount of communication channels used, but it decreased the computational costs by a factor n.\\
A second big strategy used by SPDZ and Overdrive is the so called Soundness Slack. Because in practice it is to expensive to prove that a result is exact, so it is sufficient to prove that the value, here the product of the beaver triple, exact but with the legalization of a small error. Later on, this error, if existing and introduced by an adversary, will be detected by a different technique called sascrificing.\\
Overdrive uses as well drowning, which is the technique of adding noise to a secret sharing. Because otherwise, the noise of the SHE function would reveal information over the secret. So the noise is drowned by introducing an additional random noise that is bigger than the maximum possible noise.\\

\subsubsection{TopGear}
Recently, an even newer improvement for the offline phase was introduced, called TopGear \cite{cryptoeprint:2019:035}. It is implemented in SCALE-MAMBA since version 1.3\cite{ScaleMambaDocu}, which was released in January 2019. It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs.

\subsubsection{Online Phase}

\subsubsection{Preliminaries}

Now we assume that the offline phase was succesfully executed, so all preliminaries are given to execute the online phase. It's main goal is to evaluate a specified circuit with secret inputs by the different parties. So this is the heart of the system, the part where the real computatiuon is executed. 
It is important to note, that the online phase is nearly perfectly optimized when assuming the offline phase was carried out succesfully. The online phase has a runtime complexity in O(n), and therefore there is not a really big performance impact. It is also noteworthy, that in the online phase only only computational and no information theoretical problems are solved. All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ ~\cite{cryptoeprint:2011:535}.\\

We assume following preliminaries are given after the execution of the offline phase:\\
\begin{enumerate}
\item There are \(n\) parties \(P_1,...,P_n\)
\item There is a given function \(f(a_1,...,a_m)\) which the parties want to evaluate together. The inputs are secret, and each input is hold by exactly one party.
\item There is a global MAC key \( \alpha\) that is secret to every party and additively shared, so that every party \( P_i\) has a share  \( \alpha_i\) so that  \( \alpha=\alpha_1+...+\alpha_n\)
\item Each of the secret inputs to \(f\) is additively shared between the parties

\end{enumerate}

Following, it is going to be demonstrated, how addition and multiplication are done in this environment. This is sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation.\\

Sharing of values:\\
After sharing a value \( a \in F_p\), it is represented as following tuple:\\
$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
\\With:
\begin{itemize}
\item $ a=(a_1+...+a_n)$
\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)

\end{itemize}

\subsubsection{Addition}
Assumed there are two values \( a\) and \( b\) which should be added together and both are secret inputs from two different parties. Naturally, both of these values are additively shared between all the parties, i.e. every party has a share \( \langle a_i\rangle\) and a share \( \langle b_i\rangle\).\\

Then every party \( P_i\) adds up their local shares, so that:\\
$$\langle c_i\rangle=\langle a_i\rangle+\langle b_i\rangle=((\delta_a+\delta_b),(a_i+b_i),\gamma (a)_i+\gamma (b)_i))$$
\textbf{Correctness}: Afterwards, \( c=a+b\) is obviously additively shared between the parties and can be used for further computations.\\

\subsubsection{Addition with constant}
Adding a public constant is really simple in this representation. Because \( \gamma (a)=\alpha(a+\delta)\), adding a constant \(c\) to a additively shared value \(a\) is done by:\\
$$\langle a\rangle+c=(\delta-c,(a_1+c,a_2,...,a_n),,(\gamma (a)_1,...,\gamma (a)_n)$$
Correctness: It can be seen, that the constant is only added to one shared value and subtracted from \( \delta\). Because of that, \( \alpha(a+\delta)\) stays the same, so the same shared MAC values like before can be used. So \(a+c\) is also properly additively shared.\\

\subsubsection{Multiplication with constant}
Multiplicating with a public constant is as well really simple and can also be done by each party invdividually. It is defined as:\\
$$c*\langle a\rangle=\langle c*a\rangle=(c*\delta,(c*a_1,...,c*a_n),,(c*\gamma (a)_1,...,c*\gamma (a)_n)$$
\textbf{Correctness}:(TO BE DONE\\

\subsubsection{Multiplication}

Multiplication is the only operation,except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
Assumed there are two values \( a\) and \( b\) which should be multipled with each other. Naturally, each party has the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
For each multiplication operation, we need one of the Beaver triples \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) generated in the offline phase. In the offline phase we checked the quality of our triples, but we still allowed the possibility that each triple has an error \(e\) so that \(c=a*b+e\). To check the quality of our triple, we use a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing it. Both triples are additively shared between all parties like the input values. In SPDZ both triples are completly unrelated to each other, and the sacrificing is done as follows:\\

\begin{enumerate}
\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n)\).
\item The parties calculate  \( p=t*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
\item The parties calculate \(\phi = t * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p)\) and then partially open the result
\item If \( \phi \) is 0 then the triples are ok and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\). Otherwise the system detected a malicious adversary and it aborts.\\
\end{enumerate}

Partially opening  of values: All players send their respective shares of a value, but not the MAC value, to one chosen party \(P_i\) which summs up all values and then broadcasts the result. In praxis this party is changing for every opening to balance the workload.\\
Fully opening of values: In this case all players broadcast their respective shares and the respective MACs to all players, so all players can verify the result.\\

\textbf{Correctness}: If we simplify \( \phi \), we can see that every term disappears (given that the triples are proper triples):
$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
So we can see that \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so we know the triple is correct. \\
In this process we didn't reveal the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \), so now we can use it for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\

There is a small improvement in SCALE-MAMBA compared to SPDZ regarding the triple generation that was added later in ~\cite{cryptoeprint:2017:123}. The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similiar triple. For this, the triples in the online phase are already generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle x*y \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle x'*y\rangle ) \). This way the triples can be calculated faster in the offline phase, because one half of the multiplication is the same, and still the triples keep the property to be sacrificed for each other to check the correctness of the other triple.\\\\

In SCALE-MAMBA, this triple sacrifice is also already executed in the offline phase. But for consistency the explanation remains here, to better be able to see the difference between SPDZ and SCALE-MAMBA.\\

Now in the next step the multiplication of \(\langle a \rangle \) and \(\langle b \rangle \) is executed with the help of the triple.
\begin{enumerate}
\item The parties calculate \( \epsilon = \langle a \rangle - \langle x_1 \rangle \) and partially open it
\item The parties calculate \( \delta = \langle b \rangle - \langle y_1 \rangle \) and partially open it
\item The parties calculate \( \langle c \rangle \langle z_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
\end{enumerate}
Correctness: 
$$  \langle a \rangle * \langle b \rangle = \langle (a-x_1+x_1) \rangle * \langle (b-y_1+y_1) \rangle$$
$$ = ( \epsilon  +\langle  x_1 \rangle ) * ( \delta  + \langle y_1 \rangle )$$
$$ = \langle x_1 \rangle * \langle y_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta =  \langle c \rangle $$

So, as we can see, the calculations are correct and at the same time no information of \(\langle a \rangle \) or \(\langle b \rangle \) was revealed.\\

\subsubsection{Verifying Correctness}
At the end of the computation, we have to find out the total calculated result. After all computations were executed, every party has a share of the final result \( r_i\), a share of the MAC value \( \gamma(r)_i)\) of the final result and a share of the global MAC key \( \alpha_i \). The verification is done as follows:

\begin{enumerate}
\item Every party  opens their share \( r_i\) and their share of the MAC value \( \gamma(r)_i) \)
\item Every party computes \( r = \sum_{i=1}^n r_i \) 
\item Every party computes  \( \gamma(r)_i = \sum_{i=1}^n \gamma(r)_i \)
\item Every party computes  \( \sigma_i =  \gamma(r)_i - \alpha_i * r\) and shares it to all the other parties
\item Every party computes \( \sum_{i=1}^n \sigma_i \) and checks if it is \(0\). If it is  \(0\), the computation was correct.
\end{enumerate}
Correctness: 
$$ \sum_{i=1}^n \sigma_i =  \sum_{i=1}^n  \gamma(r)_i - \alpha_i * r $$
$$=  \sum_{i=1}^n  \gamma(r)_i -  \sum_{i=1}^n  \alpha_i * r $$

If the calculations were correct, the MAC was shared properly and not modified, and the MAC values were not modified, then the difference between both sums will be 0. If it is not 0, the protocol detects that there is a malicious adversary and aborts.\\

\subsubsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time. That would offere the possibility, that the offline data is calculated before and then, when the need arises, a computatation could be specified in MAMBA and then intime compiled and executed, without having to wait for the expensive offline calculations.\\

\section{Comparison}\label{sec:comparison}

a.	What else is out there?
b.	Different use cases – when which SMPC tool is more useful; what are the strengths of SCALE-MAMBA compared to other ones?
	- Security! High performant for such a secure system
	- In some cases not the highest security is needed; application for a normal user, not for a company, have much less security requirements but much higher requirements on performance
	- Active security is not always required, in non-critical systems it is often sufficient to know that noone knows the users data
c.	Factual comparisons
	- YAO/LSSS --> LSSS protocols right now are more suited for the praxis

	- SCAPI
	-SCAPI is an open-source Java library for implementing secure two-party and multiparty computation protocols (SCAPI stands for the “Secure Computation API”). It provides a reliable, 	efficient, and highly flexible cryptographic infrastructure.
SCAPI provides only passive security

https://scapi.readthedocs.io/en/latest/intro.html

\section{Resume}\label{sec:outlook}

\subsubsection{Outlook}
Their are a lot of future applications for Secure Multiparty Computation. With the threoretical research of the early years much more turning to praxis oriented research, todays SMPC protocols, and especially the successors of SPDZ, have a bright future. With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, the list will only grow more and more with the performance of the protocols and the same time the available computation power improving. \\
With SCALE-MAMBA being one of the first performant systems that can be really used in a real time scenario, it will probably not take a lot of more years until there will be proper commercial solutions. And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distrbuted environments. There are a lot more possible applications that could be coming in the next years, for example in the financial branche SMPC could play a huge role.\\
SMPC can be a technology that shapes the future, or it can be just a stepstone for a new technology to emerge. But definetly, privacy and security plays already a huge role in our society, and this will not diminish in any close future.

\subsubsection{Conclusion}

In this paper we could see how SMPC works in general, and, especially, how SCALE-MAMBA works and how it evolved over the years in comparison to its predecessor, SPDZ. It was shown, how the protocol is split into a primary offline phase in which the computational expensive calculations are executed and all the preparations are done for the later circuit evaluation, like calculating the Beaver Triples, and how afterwards in the online phase the circuit is evaluated, how the triples are used to execute a multiplication without sharing the secret inputs and how the results can be verified at the end.

\renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
% Use natbib compatbile splncsnat style.
% It does provide all features of splncs03, but is developed in a clean way.
% Source: http://phaseportrait.blogspot.de/2011/02/natbib-compatible-bibtex-style-bst-file.html
\bibliographystyle{splncsnat}
\begingroup
  \ifluatex
    %try to activate if bibliography looks ugly
    %\sloppy
  \else
    \microtypecontext{expansion=sloppy}
  \fi
  \small % ensure correct font size for the bibliography
  \bibliography{paper}
\endgroup

% Enfore empty line after bibliography
\ \\
%

\end{document}